---
title: "Hypothesis Testing"
author: "Etana Disasa"
date: "11/15/2018"
output: pdf_document
---

## EXERCISE
```{r exercise}
a_training = c(12.9, 13.5, 12.8, 15.6, 17.2, 19.2, 12.6, 15.3, 14.4, 11.3)
b_training = c(12.7, 13.6, 12.0, 15.2, 16.8, 20.0, 12.0, 15.9, 16.0, 11.1)
t.test(a_training, b_training, paired=TRUE)
qt(0.95, 9)
```
## In the above hypothesis testing, the p-value is 0.8358 which is greater than 0.05. Therefore, the new coach's training had not made any significant change to the team's performance. 

## The t-value generated by the test is less than t-tablulated therefore, we accept the null hypothesis. 

```{r}
a_training = c(12.9, 13.5, 12.8, 15.6, 17.2, 19.2, 12.6, 15.3, 14.4, 11.3)
b_training = c(12.0, 12.2, 11.2, 13.0, 15.0, 15.8, 12.2, 13.4, 12.9, 11.0)
t.test(a_training, b_training, paired=TRUE, alt="less")
pt(0.95, 9)
```
## However, in this second coach's training, the p-value is larger than 0.05. Furthermore, given the alt="less" being the hypothesis, we reject the null hypothesis which means that the new coach's training has caused for a change. 


## DRUG TESTS XYX and ABC
## Hypothesis is testing weather xyz or abc affect the BP of a patient using either. 
```{r drug_test, message=FALSE}
library(readxl)
library(data.table)
drug.trial <- read_excel("paired-t-test_3.xlsx")
before.trial <- melt(drug.trial[4,], value.name = "before.trial")$before.trial
after.trial.xyz <- melt(drug.trial[5,], value.name = "after.trial.xyz")$after.trial.xyz
after.trial.abc <- melt(drug.trial[11,], value.name = "after.trial.abc")$after.trial.abc

trial <-cbind(before.trial, after.trial.xyz, after.trial.abc)

t.test(trial[,1], trial[,2], paired=TRUE)
```
## Here, the p-value is greater than 0.05. Therefore, there is not a signficant difference in patient's BP after using XYZ. 

```{r}
t.test(trial[,1], trial[,3], paired=TRUE)
```

## However, there, the p-value is significanly less than 0.05. Therefore, we reject the null hypothesis and take the hypothesis.  Thus, there is a signficant difference and change in BP after patient's use medication abc. 