---
title: 'ASSIGNMENT TWO: EDA'
author: "Etana Disasa"
date: "11/8/2018"
output: pdf_document
---

```{r setup, include=FALSE}

library(ggplot2)
library(data.table)
library(e1071)
library(stringr)
library(dplyr)
library(ggpubr)
```

## Factorial Design

This approach takes more than factors to consider to do the experiment. 
```{r}
df3 = read.csv("fastfood-3.csv") 
r = c(t(as.matrix(df3))) # response data 
f1 = c("Item1", "Item2", "Item3") # 1st factor levels 
f2 = c("East", "West")            # 2nd factor levels 
k1 = length(f1)          # number of 1st factors 
k2 = length(f2)          # number of 2nd factors 
n = 4  
tm1 = gl(k1, 1, n*k1*k2, factor(f1))
tm1
tm2 = gl(k2, n*k1, n*k1*k2, factor(f2)) 
tm2

av = aov(r ~ tm1 * tm2) 
summary(av) 
```

## Randomized Block Design
Here, this design only takes one primary factor to consider for the experimant. Each variable is tested against all treatments. 
```{r}
df2 = read.table("fastfood-2.txt", header=TRUE); df2 
r = c(t(as.matrix(df2))) # response data 
r 
f = c("Item1", "Item2", "Item3")   # treatment levels 
k = 3                    # number of treatment levels 
n = 6                    # number of control blocks


tm = gl(k, 1, n*k, factor(f))   # matching treatment 
tm 

blk = gl(n, k, k*n)             # blocking factor 
blk 

av = aov(r ~ tm + blk)

summary(av) 

```

## Completely Randomaized Design
Similarly, there is only one primary factor to consider; however, the test subjects are assigned to treament levels of the primary factor at random. 
```{r}
df1 = read.table("fastfood-1.txt", header=TRUE); df1 


r = c(t(as.matrix(df1))) # response data 
r 

f = c("Item1", "Item2", "Item3")   # treatment levels 
k = 3                    # number of treatment levels 
n = 6                    # observations per treatment

tm = gl(k, 1, n*k, factor(f))   # matching treatments 
tm 

av = aov(r ~ tm)

summary(av) 
```

